\texttt{Taxonomic supertree construction with}\texttt{\emph{ Incertae sedis}}\texttt{ taxa}

### 2018-05-07 doc/incertae_sedis.tex
up to listed cases

Specifically, the methods described here only consider cases in which 
    $\mathcal{P}_n$ can be determined from the structure of the tree.
Our method takes $\mathcal{P}_n$ to be the set of nodes that
    are in the taxonomic subtrees rooted at the sibling taxa of \emph{incertae sedis} taxon $n$.
This allows us to encode the exclude set of some node $x$ efficiently by sweeping
    over the tree from root to tip.
If $\mathcal{S}(x)$ is the set of siblings of $x$ which are not annotated as being 
    \emph{incertae sedis} and $p(x)$ is the parent of $x$, then
\begin{align}
    \mathcal{E}(x) = \mathcal{E}(p(x)) \cup \bigcup_{s\in \mathcal{S}(x)} \mathcal{E}(s)
\end{align}
with $\mathcal{E}(\rho) = \emptyset$ for the root of the taxonomic tree.

@@

\caption{\label{fig:Broken-incertae-sedis} \emph{Incertae sedis} clade $T$
broken by conflicting placement.
Input trees (a) and (b) conflict
in the placement of taxa in $T$.
The synthesis supertree (d) places
places $T1$ and $T2$ separately within $R$, while unplaced taxa
float upwards.
}

@@

In order to examine the results of the rooted splits-semantics above, we explore
the outcome on a series of complex cases.  These cases are intended to
illustrate some of the issues that arise when developing an operational semantics of
\emph{incertae sedis}, as well as the outcomes for the rooted-split semantics.
We think that they constitute one of the main contributions of the paper.  In
each case, the synthesis trees in this section are constructed using the
BUILD-based algorithm described in section \ref{sec:background}, without
decomposing them into smaller subproblems or pruning and then unpruning
taxonomy-only taxa.  Synthesis trees that are described as \emph{incertae
sedis}-aware apply the BUILD algorithm to splits constructed using equation
\ref{eq:exsetformonea}, whereas the synthesis trees that are not \emph{incertae
sedis}-aware are constructed from splits using equation
\ref{eq:exsetformonea-traditional}.

@@

With the naive semantics, intrusion of an incertae sedis taxon into a sibling is
interpreted as conflict with that sibling.  This results in the loss of the edge
corresponding to the sibling taxon as well as the taxon name.

We see from Fig. \ref{fig:Handling-incertae-sedis}d that use of the incertae sedis
semantics results the taxonomy edge $AB$ grouping $A$ and $B$.  It also restores the
names $AB$ and $A$.  We may thus determine that $C$ has been placed within $A$ (and
its ancestor $AB$).

\subsection{Conflict with incertae sedis taxa}

>>>>

Specifically, the methods described here only consider cases in which 
    $\mathcal{P}_n$ can be determined from the structure of the tree.
Our method takes $\mathcal{P}_n$ to be the set of nodes that
    are in the taxonomic subtrees rooted at the sibling taxa of \emph{incertae sedis} taxon $n$.
This allows us to encode the exclude set of some node $x$ efficiently by sweeping
    over the tree from root to tip.
If $\mathcal{S}(x)$ is the set of siblings of $x$and $p(x)$ is the parent of $x$, then
the naive approach of ignoring the \emph{incertae sedis}  annotation amount to setting
    the exclude set as:
\begin{align}
    \mathcal{E}(x) = \mathcal{E}(p(x)) \cup \left\{ \bigcup_{s\in \mathcal{S}(x)} \mathcal{I}(s) \right\}
    \label{eq:exsetformonea-traditional}
\end{align}
with $\mathcal{E}(\rho) = \emptyset$ for the root of the taxonomic tree.

If we treat the  \emph{incertae sedis} annotation as an indication that the 
    \emph{incertae sedis} taxon $x$ can be place in any of its sister taxa, then
    we let $\mathcal{Q}(x)$ denote the set of siblings of $x$ that are not
    flagged as being \emph{incertae sedis} and we define the exclude set as:
\begin{align}
    \mathcal{E}(x) = \mathcal{E}(p(x)) \cup \left\{ \bigcup_{s\in \mathcal{Q}(x)} \mathcal{I}(s)\right\}
    \label{eq:exsetformonea}
\end{align}
We will refer to this interpretation of the \emph{incertae sedis} annotation as ``unrestricted
    intrusion'' semantics.

@@

\caption{\label{fig:Broken-incertae-sedis} \emph{Incertae sedis} clade $T$
broken by conflicting placement.
Input trees (a) and (b) conflict
in the placement of taxa in $T$.
The synthesis supertree (d) places
places $T1$ and $T2$ separately within $R$.
The BUILD algorithm would place the unsampled taxon $T3$ at the
root of the tree, while our unprune operation implemented to 
handle taxonomy-only taxa would place $T3$ as a child of the MRCA
of the other members of $T$, as pictured in (d).
}

@@

In order to examine the results of the rooted splits-semantics above, we explore
the outcome on a series of complex cases.  These cases are intended to
illustrate some of the issues that arise when developing an operational semantics of
\emph{incertae sedis}, as well as the outcomes for the rooted-split semantics.
%We think that they constitute one of the main contributions of the paper.
In each case, the synthesis trees in this section are constructed using the
BUILD-based algorithm described in section \ref{sec:background}, without
decomposing them into smaller subproblems or pruning and then unpruning
taxonomy-only taxa.  Synthesis trees that are described as \emph{incertae
sedis}-aware apply the BUILD algorithm to splits constructed using equation
\ref{eq:exsetformonea}, whereas the synthesis trees that are not \emph{incertae
sedis}-aware are constructed from splits using equation
\ref{eq:exsetformonea-traditional}.

@@

With the naive semantics, intrusion of an \emph{incertae sedis} taxon into a sibling is
interpreted as conflict with that sibling.  This results in the loss of the edge
corresponding to the sibling taxon as well as the taxon name.

We see from Fig. \ref{fig:Handling-incertae-sedis}d that use of the unrestricted intrusion
semantics results the taxonomy edge $AB$ grouping $A$ and $B$.  It also restores the
names $AB$ and $A$.  We may thus determine that $C$ has been placed within $A$ (and
its ancestor $AB$).

\subsection{Conflict with \emph{incertae sedis} taxa}
